---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
First step will be to load the required package and download the data directly in the folder.

```{r}

  require(dplyr)

  if(!file.exists("PA")){
    dir.create("PA")
  }


    fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"


    download.file(fileurl,destfile="./PA/PA.zip")
    filelist<-unzip("./PA/PA.zip")

    date_download<-date()



    f1<-tbl_df(read.csv(filelist))

```

Next is to inspect the acquired data.Turns out the table consists of 3 variables.  
1. Steps  
2. Date  
3. Interval

```{r}
    
    head(f1)
    str(f1)
    summary(f1)

```



## What is mean total number of steps taken per day?


```{r}
  dat<-group_by(filter(f1,!is.na(steps)),date)
    
  dat_sum<- summarize(dat,sum(steps))
  
  dat_mean<-round(mean(dat_sum$`sum(steps)`))
  dat_median<-median(dat_sum$`sum(steps)`)
  
  
  hist(dat_sum$`sum(steps)`, breaks= 20, main = 'Histogram',xlab = 'Total Number of Steps')
  
  abline(v=dat_mean, lwd = 1, col = 'blue')
  abline(v=dat_median, lwd = 1, col = 'red')
  
```

Upon calculating the mean and median.

Mean = `r dat_mean`
Median = `r dat_median`

Their values somewhat conincide.
  
  

## What is the average daily activity pattern?


Now we shall construct a timeseries plot, where we will visulise the number of steps taken throughout the time interval


```{r}
  dat<-group_by(filter(f1,!is.na(steps)),interval)

  dat_mean<- summarize(dat,mean(steps))
  
  max_v<-dat_mean[which.max(dat_mean$`mean(steps)`),]
  max_v_step = max_v$`mean(steps)`
  
  
  plot(dat_mean$interval,dat_mean$`mean(steps)`,type = 'l',main = 'Average Steps by Time Interval',xlab     ='5 Minute Time Interval',ylab = 'Average Number of Steps')
  
   points(max_v$interval,max_v$`mean(steps)`, col = 'red', lwd = 3, pch = 19)
        
```

Maximum value at `r max_v$interval` is `r max_v_step`

## Imputing missing values
We will impute the missing value using the Mice (multiple imputation by chained equation) function from mice package.

You need to create a dataframe with the variable you want to impute, and include every variable that might predict values of that variable (so every var. in your model + possibly other var. as well). The mice package will impute every missing value in that dataframe.

```{r simulation, results= "hide"}
library(mice)
imputed_data<-mice(f1, m = 2, maxit = 50, method = 'pmm', seed = 500)
imputed_data$imp$steps
complete_Data<-complete(imputed_data,2)


```

```{r}
head(complete_Data)

complete_Data <- group_by(complete_Data, date)
complete_sum<- summarize(complete_Data,sum(steps))

hist(complete_sum$`sum(steps)`, breaks= 20, main = 'Histogram',xlab = 'Total Number of Steps')

completed_mean<-round(mean(complete_sum$`sum(steps)`))
completed_median<-median(complete_sum$`sum(steps)`)

abline(v=completed_mean, lwd = 1, col = 'blue')
abline(v=completed_median, lwd = 1, col = 'red')
legend("topright",pch = 1, col = c("blue","red"),legend = c("Mean","Median"))

```

## Are there differences in activity patterns between weekdays and weekends?
We will filter out Weekends and Weekdays using as.Date function and then plotting the Average steps against interval


```{r}

complete_Data$days<-weekdays(as.Date(complete_Data$date))

weekends<-filter(complete_Data,days == c("Saturday","Sunday"))
weekdays<-filter(complete_Data,days != c("Saturday","Sunday"))

weekends<-group_by(weekends,interval)
weekdays<-group_by(weekdays,interval)

weekends_mean<- summarize(weekends,mean(steps))
weekdays_mean<-summarize(weekdays,mean(steps))

par(mfrow = c(1, 2))
plot(weekends_mean$interval,weekends_mean$`mean(steps)`,type = 'l',main = 'Average Steps by Time Interval in Weekend',xlab ='5 Minute Time Interval',ylab = 'Average Number of Steps')

plot(weekdays_mean$interval,weekdays_mean$`mean(steps)`,type = 'l',main = 'Average Steps by Time Interval in Weekday',xlab ='5 Minute Time Interval',ylab = 'Average Number of Steps')

```

